\section{Blockchain Construction}\label{sec:blockchain_construction}

The blocks of the blockchain contain information used for the proof
of space, called \emph{trunk} by borrowing this terminology from~\cite{CohenP19},
other information such as the previous block hash,
and extra information that is irrelevant here, such as a list of transactions.
Transactions are not formalized below, since they are not used for the proof of space.
Blocks can be genesis and non-genesis. Both contain their time of creation and their acceleration.
Genesis blocks have nor trunk nor parent and their height is implicitly $0$.
Challenges $c$ are generated in sequence: there is an initial constant challenge for the genesis
blocks, while the subsequent challenges are generated from the trunk of
each non-genesis block $b$ of the blockchain.
In particular, $c$ is \emph{not} computed from the transactions in $b$,
in order to avoid block-grinding attacks (Sec.~\ref{sec:related_work}).
A deadline that answers $c$ is recorded in the trunk of the sons of $b$.
%
\begin{definition}[Trunk, Block]\label{def:trunk}
  The sets of \emph{trunks} and \emph{blocks} are
  \[
  \Trunks=\left\{\langle\height,\delta\rangle\left|\;\height\in\mathbb{N}\text{ and }\delta\in\Deadlines\right.\right\}
  \]
  \[
  \GenesisBlocks=\left\{\langle\tau,\alpha\rangle\mid\tau\in\mathbb{N},\ \alpha\in\mathbb{N}\text{ and }\alpha>0\right\}
  \]
  \[
  \NonGenesisBlocks=\left\{\left\langle\begin{array}{c}
  \tau,\alpha,\\
  \power,\weightedbeat,\\
  \trunk,\\
  \previousblockhash
  \end{array}\right\rangle\left|\begin{array}{l}
  \tau\in\mathbb{N},\\
  \alpha\in\mathbb{N},\ \alpha>0,\\
  \power\in\mathbb{N},\\
  \weightedbeat\in\mathbb{N},\\
  \trunk\in\Trunks,\\
  \previousblockhash\text{ is a}\\
  \qquad\text{hash for $h_\block$}
  \end{array}\right.\right\}
  \]
  \[
  \Blocks=\GenesisBlocks\cup\NonGenesisBlocks.
  \]
\end{definition}
%
If $b$ is a block, then $b.\tau$ is its creation time (milliseconds from the Unix epoch)
and $b.\alpha$ is the acceleration at $b$. If $b$ is a non-genesis block, then
$b.\power$ expresses how much space has been used to build the path that leads to $b$,
starting from the genesis block; it will be used to select the \emph{best chain} for mining
$b$'s son. The value of $b.\weightedbeat$ is the average of the interval creation time
for the path that leads to $b$, giving more weight to the last blocks; it will be
compared to $\beat$ (Tab.~\ref{tab:notations}) to understand if the acceleration
must be increased or decreased in $b$'s son.
The value of $b.\previousblockhash$ is the hash of the previous block in the path leading to $b$.
If $b$ is a genesis block, we abuse notation and assume that $b.\power=b.\weightedbeat=0$.

\begin{definition}[Block's height]\label{def:block_height}
  Let $b\in\Blocks$. The \emph{height} of $b$ is defined as
  \[
  \height(b)=\begin{cases}
  0 & \text{if $b\in\GenesisBlocks$}\\
  b.\trunk.\height & \text{if $b\in\NonGenesisBlocks$.}
  \end{cases}
  \]
\end{definition}

The following algorithm shows how the first challenge, for genesis blocks,
its defined. It is a constant that only depends on contextual values (Table~\ref{tab:notations}).
%
\begin{alg}[$\initialchallenge$]\label{alg:initial_challenge}
  The \emph{initial challenge} is the constant
  %
  \begin{multline*}
    \initialchallenge=\\
    \langle\betonat(h_\generation(\sigma_\genesis\append\nattobe(1)))\text{ mod }\numberofscoops,\sigma_\genesis\rangle,
  \end{multline*}
  %
  where $\sigma_\genesis$ is a constant generation signature used for the genesis of the blockchain
  (see Table~\ref{tab:notations}).
\end{alg}
%
The following algorithm shows how a challenge is derived from the trunk of a non-genesis block.
%
\begin{alg}[$\nextchallenge(\trunk)$]\label{alg:next_challenge_from_trunk}
  Let $\trunk\in\Trunks$. The \emph{next challenge for $\trunk$} is
  \begin{multline*}
    \nextchallenge(\trunk)=\\
    \langle\betonat(h_\generation(\sigma\append\nattobe(\trunk.\height+1)))\text{ mod }\numberofscoops,\sigma\rangle
  \end{multline*}
  where
  \[
  \sigma=h_\generation(\trunk.\delta.\challenge.\sigma\append\trunk.\delta.\pi).  
  \]
\end{alg}
%
The definition of the generation signature $\sigma$ for the next challenge,
in Alg.~\ref{alg:next_challenge_from_trunk},
has puzzled us for some time, since~\cite{SignumPlotting} appends a \emph{previous block generator}
to the previous block's generation signature $\trunk.\delta.\challenge.\sigma$.
That concept, however, is defined nowhere.
We had to dive in the source code of the Signum node
to understand that it is actually
an identifier (more concretely, the public key)
of the creator of the deadline for the previous block
(see \url{https://github.com/signum-network/signum-node/blob/main/src/brs/GeneratorImpl.java}, constructor of \<GeneratorStateImpl>).
We abstract that information through the prolog of the previous deadline, hence this is why
Alg.~\ref{alg:next_challenge_from_trunk} appends
$\trunk.\delta.\pi$ to define $\sigma$.
This gives us the opportunity to discuss the observation in~\cite{ParkPAFG15}, page~28,
where it is stated that ``this previous block generator seems possible to be grinded, by trying
different sets of transactions to include in a block''. This is false, although it was
hard to grasp without the formalization in
Alg.~\ref{alg:next_challenge_from_trunk}, where it is clear that
$\trunk.\delta.\pi$ comes from the trunk only, which makes
such grinding attack impossible.

Later, it will be handy to determine the next challenge for a block. Note that
the next definition only uses the trunk inside the block.
%
\begin{definition}\label{def:next_challenge_from_block}
  Let $b\in\Blocks$. Its \emph{next challenge} is
  \[
  \nextchallenge(b)=\begin{cases}
  \initialchallenge & \text{if $b\in\GenesisBlocks$}\\
  \nextchallenge(b.\trunk) & \text{if $b\in\NonGenesisBlocks$.}
  \end{cases}
  \]
\end{definition}

In the following, it is shown how the information inside a block is used to construct
the information inside its son(s). The computation of the next weighted beat
gives more or less weight to the previous weighted beat, depending on a constant $\oblivion$.
Therefore, the latter ($0\le\oblivion\le 1$) expresses how quickly the acceleration reacts to changes
in mining power. The computation of the next power uses the same formula of Bitcoin~\cite{WalkerG24},
adapted to our context: the maximal (hence worse) deadline's value $2^{8\cdot\size(h_\deadline)}$ is compared
to the actual deadline's value and their ratio expresses how much space has been used to compute the deadline.
%
\begin{definition}[Next functions]\label{def:next}
  Let $b\in\Blocks$ and $\delta\in\Deadlines$. We define
  \[
  \nextwaitingtime(b,\delta)=b.\tau+\waitingtime(\delta,b.\alpha)
  \]
  \[
  \nextweightedbeat(b,\delta)=\begin{array}{c}
  \waitingtime(\delta,b.\alpha)\cdot\oblivion\\
  +b.\weightedbeat\cdot(1-\oblivion)
  \end{array}
  \]
  \[
  \nextacceleration(b,\delta)=\frac{b.\alpha\cdot\nextweightedbeat(b,\delta)}{\beat}
  \]
  \[
  \nextpower(b,\delta)=b.\power+\frac{2^{8\cdot\size(h_\deadline)}}{\betonat(\delta.\mathit{value})+1}
  \]
\end{definition}
%
\begin{definition}[Next block]\label{def:next_block}
  Let $b\in\Blocks$ and $\delta\in\Deadlines$. We define
  $\nextblock(b,\delta)\in\NonGenesisBlocks$ as
  \[
  \nextblock(b,\delta)=\left\langle\begin{array}{c}
  \nextwaitingtime(b,\delta),\nextacceleration(b,\delta),\\
  \nextpower(b,\delta),\nextweightedbeat(b,\delta),\\
  \langle\height(b)+1,\delta\rangle,\\
  h_\block(b)
  \end{array}\right\rangle.
  \]
\end{definition}

A blockchain is a set of blocks, linked through their $\previousblockhash$ field.
It must contain exactly one genesis block; it has no hash collisions among its blocks;
and all its blocks must satisfy the \emph{consensus rules}.
%
\begin{definition}[Blockchain, Consensus]\label{def:blockchain}
  A \emph{blockchain} is a set $B\subset\Blocks$ such that:
  \begin{enumerate}
  \item there is exactly one $b\in B\cap\GenesisBlocks$, written as $\genesis(B)$;
  \item for each hash $h$ of $h_\block$, there is at most
    one $b\in B$ such that $h_\block(b)=h$, written as $\block(B,h)$;
  \item for each $b\in B$, the predicate $\consensus(B,b)$ holds, where
    %
    \begin{itemize}
    \item if $b\in\GenesisBlocks$, then the only requirement for consensus is that
      $b$ is not created in the future:
      \[
      \consensus(B,b)=b.\tau\ge\tau_\now;
      \]
    \item if $b\in\NonGenesisBlocks$, then $\consensus(B,b)$ is the logical conjunction
      of all the following consensus rules:
      \begin{itemize}
      \item $b$ is not created in the future:
        \[
        b.\tau\ge\tau_\now;
        \]
      \item the deadline of $b$ (that is, $b.\trunk.\delta$) is valid (Def.~\ref{def:deadline_validity});
      \item there are no dangling pointers:
        \[
        p=\block(B,b.\previousblockhash)\text{ exists;}
        \]
      \item the deadline of $b$ answers the challenge of $p$ (Def.~\ref{def:next_challenge_from_block}):
        \[
        \nextchallenge(p)=b.\trunk.\delta.\challenge;
        \]
      \item $b$ is the next block of $p$ \wrt the deadline of $b$ (Def.~\ref{def:next_block}):
        \[
        b=\nextblock(p,b.\trunk.\delta).
        \]
      \end{itemize}
    \end{itemize}
    %
  \end{enumerate}
\end{definition}
%
Note that the above consensus rules, reconstructed and interpolated
from~\cite{SignumPlotting,SignumSource},
do not constrain the prolog of the deadlines in any way:
each block can have an arbitrary prolog. Later, it will be shown why it is useful to
restrain prologs with an extra consensus rule.

\begin{definition}[Blockchain network]\label{def:blockchain_network}
  A \emph{blockchain network} is a
  network of \emph{peers} (computers), each connected to the other peers,
  each holding its own version of a blockchain, for the same genesis block,
  not created in the future.
  Each peer holds a plot (Def.~\ref{def:plot}) in its memory.
  Each peer starts with a blockchain holding only the genesis block and runs two
  algorithms, concurrently: the \emph{block mining} algorithm
  and the \emph{block mined} algorithm.
\end{definition}
%
Note that Def.~\ref{def:blockchain_network} simplifies the picture very much:
the peers are fully connected, never disconnect and never need to synchronize.
Moreover, in practice, peers do not hold plots but rather rely on (one or more)
external services (miners) that hold one or more plots.
The goal here is to keep the picture as simple as possible and concentrate on the properties of the
proof of space algorithm only. Def.~\ref{def:blockchain_network} does not pretend
to describe a real blockchain implementation.

The block mining algorithm looks for the most powerful block in blockchain,
mines a new block on top of it, adds it to the blockchain and whispers it to the other peers.
%
\begin{alg}[Block mining]\label{alg:block_mining}
  The \emph{block mining} algorithm of a peer $P$, holding blockchain $B$,
  is the following infinite loop:
  %
  \begin{enumerate}
  \item identify a most powerful\footnote{In theory, more blocks might be the most powerful in blockchain, although this is highly unlikely; in that case, any of them will be chosen.} block $b$ in $B$;
  \item compute $c=\nextchallenge(b)$ (Def.~\ref{def:next_challenge_from_block});
  \item compute $\delta'=\delta(\plot,c)$ (Alg.~\ref{alg:deadline_from_plot}), where $\plot$
    if the plot of $P$;
  \item compute $b'=\nextblock(b,\delta')$ (Def.~\ref{def:next_block});
  \item\label{step:block_mining:wait} wait until $b'.\tau\ge\tau_\now$;
  \item\label{step:block_mining:add} add $b'$ to $B$;
  \item whisper $b'$ to the peers connected to $P$;
  \item go back to step~1.
  \end{enumerate}
\end{alg}
%
The block mined algorithm receives a block whispered from a peer, checks its validity
and adds it to the blockchain.
%
\begin{alg}[Block mined]\label{alg:block_mined}
  The \emph{block mined} algorithm of a peer $P$, holding blockchain $B$,
  is the following infinite loop:
  %
  \begin{enumerate}
  \item wait for a block $b$ whispered from a connected peer $P'$;
  \item\label{step:block_mined:add} if $B\cup\{b\}$ is a blockchain, add $b$ to $B$;
  \item go back to step~1.
  \end{enumerate}
\end{alg}
%
In practice, step~\ref{step:block_mined:add} of Alg.~\ref{alg:block_mined} should
allow the addition only of blocks $b$ that are \emph{powerful enough} to look useful,
in order to avoid filling the memory with useless blocks. This is not relevant in this paper.
Moreover, if the whispered block $b$
at step~\ref{step:block_mined:add} of Alg.~\ref{alg:block_mined} is more
powerful than $b'$ at step~\ref{step:block_mining:wait} of Alg.~\ref{alg:block_mining},
a rational peer would interrupt the wait at step~\ref{step:block_mining:wait}
of Alg.~\ref{alg:block_mining}, discard $b'$ and restart Alg.~\ref{alg:block_mining}
from step~1, since the whispered $b'$ is better than the block $b$ that it is trying to mine,
hence it is wiser to stop wasting time and start mining on top of the new best block $b'$.
Such optimizations are not considered here.

Note that peers check the validity of blocks coming from outside
(step~\ref{step:block_mined:add} of Alg.~\ref{alg:block_mined}), since they do not trust
their connected peers. Instead, they do not check the validity of the blocks that they
mine themselves (step~\ref{step:block_mining:add} of Alg.~\ref{alg:block_mining}),
since they are valid by construction, as shown below.
%
\begin{proposition}\label{prop:mining_is_sound}
  At step~\ref{step:block_mining:add} of Alg.~\ref{alg:block_mining},
  the set $B\cup\{b'\}$ is a blockchain.
\end{proposition}
