\documentclass[orivec]{llncs}

\pagestyle{headings}

\usepackage{showkeys}
\usepackage{hyperref}
\usepackage{url}
\usepackage{relsize}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{listings, framed}

\lstset{
  language=Java,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\ttfamily},
  frame=l,
  numbers=left,
  numberstyle={\ttfamily},
%  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
  escapechar=|
}

% \|name| or \mathid{name} denotes identifiers and slots in formulas
\def\|#1|{\mathid{#1}}
\newcommand{\mathid}[1]{\ensuremath{\mathit{#1}}}
% \<name> or \codeid{name} denotes computer code identifiers
\def\codesize{\small}
\def\<#1>{\codeid{#1}}
\newcommand{\codeid}[1]{\ifmmode{\mbox{\codesize\ttfamily{#1}}}\else{\codesize\ttfamily #1}\fi}

\newcommand{\todo}[1]{{\color{red}\bfseries [[#1]]}}

\DeclareMathOperator*{\append}{\bowtie}

\newcommand{\fs}[1]{\todo{FS: #1}}
\newcommand*\xor{\oplus}

\newtheorem{alg}{Algorithm}

\newcommand{\wrt}{\textit{wrt.\ }}
\newcommand{\ie}{, \textit{ie.}, }

\newcommand{\numberofscoops}{{\#\mathit{scoops}}}
\newcommand{\noncesize}{{\mathit{nonce}_{\mathit{size}}}}
\newcommand{\byte}{{\mathit{byte}}}
\newcommand{\nonce}{{\mathit{nonce}}}
\newcommand{\nonces}{{\mathit{nonces}}}
\newcommand{\scoops}{{\mathit{scoops}}}
\newcommand{\plot}{{\mathit{plot}}}
\newcommand{\seed}{{\mathit{seed}}}
\newcommand{\deadline}{{\mathit{deadline}}}
\newcommand{\generation}{{\mathit{generation}}}
\newcommand{\size}{{\mathit{size}}}
\newcommand{\scoopnumber}{{\mathit{scoopNumber}}}
\newcommand{\data}{{\mathit{data}}}
\newcommand{\challenge}{{\mathit{challenge}}}
\newcommand{\mynext}{{\mathit{next}}}
\newcommand{\nextchallenge}{\challenge_\mynext}
\newcommand{\initialchallenge}{{\mathit{initialChallenge}}}
\newcommand{\height}{{\mathit{height}}}
\newcommand{\trunk}{{\mathit{trunk}}}
\newcommand{\block}{{\mathit{block}}}
\newcommand{\oblivion}{{\mathit{oblivion}}}
\newcommand{\now}{{\mathit{now}}}
\newcommand{\beat}{{\mathit{beat}}}
\newcommand{\weightedbeat}{\mathit{weightedBeat}}
\newcommand{\nextweightedbeat}{\weightedbeat_\mynext}
\newcommand{\power}{{\mathit{power}}}
\newcommand{\previousblockhash}{{\mathit{previousBlockHash}}}
\newcommand{\waitingtime}{{\mathit{waitingTime}}}
\newcommand{\nextwaitingtime}{\tau_\mynext}
\newcommand{\nextpower}{\power_\mynext}
\newcommand{\nextacceleration}{\alpha_\mynext}
\newcommand{\nextblock}{\block_\mynext}
\newcommand{\genesis}{{\mathit{genesis}}}
\newcommand{\consensus}{{\mathit{consensus}}}
\newcommand{\nattobe}{{\mathit{nat2be}}}
\newcommand{\betonat}{{\mathit{be2nat}}}
\newcommand{\finalhash}{{h_{\mathit{final}}}}

\newcommand{\Prologs}{{\mathsf{Prologs}}}
\newcommand{\Plots}{{\mathsf{Plots}}}
\newcommand{\Challenges}{{\mathsf{Challenges}}}
\newcommand{\Scoops}{{\mathsf{Scoops}}}
\newcommand{\Nonces}{{\mathsf{Nonces}}}
\newcommand{\Trunks}{{\mathsf{Trunks}}}
\newcommand{\Blocks}{{\mathsf{Blocks}}}
\newcommand{\GenesisBlocks}{{\mathsf{GenesisBlocks}}}
\newcommand{\NonGenesisBlocks}{{\mathsf{NonGenesisBlocks}}}
\newcommand{\Deadlines}{{\mathsf{Deadlines}}}

%
\begin{document}
%
\begin{frontmatter}

  \title{A Formalization of Signum's Consensus}

  \author{Fausto Spoto}
  \institute{Dipartimento di Informatica, Universit\`a di Verona, Verona, Italy\\
    \email{fausto.spoto@univr.it}}

  \maketitle

  \begin{abstract}
    Bitcoin's proof of work consumes energy and currently requires
    dedicated, expensive hardware. Therefore, alternatives have been proposed,
    including proof of stake and proof of space. The latter requires to allocate
    disk space for mining, instead of CPU power. A few theoretical and
    implemented proof of space solutions exist. Signum is one of them:
    a blockchain network that runs since ten years, has a relatively simple
    consensus algorithm but lacks any formalization.
    This paper formalizes Signum's consensus algorithm and uses that formalization
    to show that Signum is free from block grinding attacks and is largely
    protected from challenge grinding attacks. Moreover, this paper proposes
    a new protection for Signum against newborn attacks.
  \end{abstract}

\end{frontmatter}

\pagenumbering{arabic}

\input{introduction}
\input{nonce_and_plot}
\input{challenge_and_deadline}
\input{blockchain_construction}
\input{prolog_and_attacks}
\input{related_work}
\input{conclusion}

\bibliographystyle{plain}
\bibliography{biblio}

\input{proofs}

\end{document}
