\newpage

\appendix
\section{Proofs (not meant for publication)}

\noindent
\textbf{Proof of Prop.~\ref{prop:mining_is_sound} at page~\pageref{prop:mining_is_sound}}

\noindent
  Let us prove it by induction on the number of blocks added to $B$ in Algs.~\ref{alg:block_mining}
  and~\ref{alg:block_mined}. The thesis is true when the peer starts,
  by Def.~\ref{def:blockchain_network}. Alg.~\ref{alg:block_mined} keeps $B$
  as a blockchain, since it explicitly checks it (step~\ref{step:block_mined:add} of
  Alg.~\ref{alg:block_mined}). It remains to show that, if $B$ is a blockchain at the beginning
  of step~\ref{step:block_mining:add} of Alg.~\ref{alg:block_mining}, then $B\cup\{b'\}$ is a
  blockchain as well. By Def.~\ref{def:next_block}, it is $b'\in\NonGenesisBlocks$.
  Therefore, property~\ref{prop:blockchain:genesis} of Def.~\ref{def:blockchain} remains true.
  By the assumption on no hash collisions, property~\ref{prop:blockchain:no_collision} remains true.
  It remains to prove that property~\ref{prop:blockchain:consensus} remains true as well.
  Let $b\in B\cup\{b'\}$.
  \begin{itemize}
  \item If $b\in\GenesisBlocks$, it is $b\not=b'$ and then $b\in B$; since $\tau_\now$ only
    increases with the time, by inductive hypothesis
    the property $b.\tau\le\tau_\now$ was true at the start of the
    algorithm and is still true now.
  \item If $b\in\NonGenesisBlocks$ and $b\in B$, by inductive hypothesis
    all consensus rules held when $b$ was added to
    $B$ and they must still hold now, since $\tau_\now$ only
    increases with the time and since no block is ever removed from $B$.
  \item If $b\in\NonGenesisBlocks$ and $b\not\in B$, it must be $b=b'$. Let us prove that each consensus rule
    in Def.~\ref{def:blockchain} holds.
    \begin{itemize}
    \item[\ref{prop:consensus:no_future}] This rule holds by step~\ref{step:block_mining:wait} of
      Alg.~\ref{alg:block_mining} and by the fact
      that $\tau_\now$ cannot decrease from step~\ref{step:block_mining:wait} to
      step~\ref{step:block_mining:add};
    \item[\ref{prop:consensus:valid}] It is
      \begin{align*}
        b.\trunk.\delta&=b'.\trunk.\delta\\
        \text{(step~\ref{step:block_mining:next_block} of Alg.~\ref{alg:block_mining} and Def.~\ref{def:next_block})}&=\delta'\\
        \text{(step~\ref{step:block_mining:next_deadline} of Alg.~\ref{alg:block_mining})}&=
        \delta(\plot,c)\\
        \text{(Def.~\ref{def:deadline_from_plot})}&=\delta(\nonce,\plot.\pi,c)
      \end{align*}
      for some $\nonce\in\plot.\nonces$. The deadline $\Delta=\delta(\nonce,\plot.\pi,c)$ is valid
      (Def.~\ref{def:deadline_validity}) since, by Def.~\ref{def:deadline_from_nonce}:
      \begin{align*}
        \Delta.\mathit{value}&=\mathit{value}(\nonce,c)\\
        &=\mathit{value}(\nonce,\Delta.\challenge)\\
        \text{(Def.~\ref{def:plot})}&=\mathit{value}(\nonce(\nonce.p,\plot.\pi),\Delta)\\
        \text{(Def.~\ref{def:deadline_from_nonce})}&=\mathit{value}(\nonce(\Delta.p,\Delta.\pi),\Delta).
      \end{align*}
    \item[\ref{prop:consensus:no_dangling}] It is
      \begin{align*}
        b.\previousblockhash&=b'.\previousblockhash\\
        \text{(step.~\ref{step:block_mining:next_block} of Alg.~\ref{alg:block_mining})}
        &=\block_\mynext(p,\delta').\previousblockhash\\
        &=h_\block(p),
      \end{align*}
      where $p$ is the most powerful block in $B$ (by step.~\ref{step:block_mining:loop}
      of Alg.~\ref{alg:block_mining}). Therefore, $\block(B,b.\previousblockhash)$ exists
      and coincides with the block $p$ selected at step~\ref{step:block_mining:loop}
      of Alg.~\ref{alg:block_mining}. This fact is used in the subsequent points of this proof.
    \item[\ref{prop:consensus:answer}] By step~\ref{step:block_mining:next_challenge}
      of Alg.~\ref{alg:block_mining}, it is
      \begin{align*}
        \challenge_\mynext(p)&=c\\
        \text{(Defs.~\ref{def:deadline_from_plot} and~\ref{def:deadline_from_nonce})}&=
        \delta(\plot,c).\challenge\\
        \text{(step~\ref{step:block_mining:next_deadline} of Alg.~\ref{alg:block_mining})}&=\delta'.\challenge\\
        \text{(step~\ref{step:block_mining:next_block} of Alg.~\ref{alg:block_mining} and Def.~\ref{def:next_block})}&=b'.\trunk.\delta.\challenge\\
        &=b.\trunk.\delta.\challenge.
      \end{align*}
    \item[\ref{prop:consensus:next_block}] It is
      \begin{align*}
        b&=b'\\
        \text{(step.~\ref{step:block_mining:next_block} of Alg.~\ref{alg:block_mining})}
        &=\block_\mynext(p,\delta')\\
        \text{(Def.~\ref{def:next_block})}&=\block_\mynext(p,b.\trunk.\delta).
      \end{align*}
    \end{itemize}
  \end{itemize}
  \qed

\noindent
\textbf{Proof of Prop.~\ref{prop:no_challenge_grinding} at page~\pageref{prop:no_challenge_grinding}}

\noindent
  Let us proceed by induction on $n$.
  If $n=0$, since $b_0=b'_0$ the thesis holds.
  Let $n>0$ and assume that the thesis holds for $n-1$. Let us prove it for $n$.
  By inductive hypothesis, it is enough
  to prove that $\challenge_\mynext(b_n)=\challenge_\mynext(b'_n)$.
  Since $n>0$, by consensus rule~\ref{prop:consensus:next_block} of Def.~\ref{def:blockchain}
  and by Def.~\ref{def:next_block}, it is $b_n\in\NonGenesisBlocks$. Therefore, by
  Def.~\ref{def:next_challenge_from_block}, it is
  %
  \begin{align*}
    \challenge_\mynext(b_n)&=\challenge_\mynext(b_n.\trunk)\\
    (Def.~\ref{def:next_challenge_from_trunk})&=\langle\betonat(h_\generation(\sigma\append\nattobe(b_n.\trunk.\height+1)))\\
    &\qquad\text{mod }\numberofscoops,\sigma\rangle\\
    \left(\begin{array}{c}
      \text{consensus}\\
      \text{rule~\ref{prop:consensus:next_block}}\\
      \text{of Def.~\ref{def:blockchain},}\\
      \text{and Def.~\ref{def:next_block}}
    \end{array}\right)&=\langle\betonat(h_\generation(\sigma\append\nattobe(b_0.\trunk.\height+n+1)))\\
    &\qquad\text{mod }\numberofscoops,\sigma\rangle
  \end{align*}
  %
  and similarly
  %
  \begin{align*}
    \challenge_\mynext(b'_n)&=\langle\betonat(h_\generation(\sigma'\append\nattobe(b_0.\trunk.\height+n+1)))\\
    &\qquad\text{mod }\numberofscoops,\sigma'\rangle
  \end{align*}
  %
  where
  %
  \begin{align*}
    \sigma&=h_\generation(b_n.\trunk.\delta.\challenge.\sigma\append b_n.\trunk.\delta.\pi)\\
    \left(\begin{array}{c}
      \text{$b_n$ has been}\\
      \text{mined by using}\\
      \text{$\plot$ only,}\\
      \text{and Def.~\ref{def:deadline_from_nonce}}
    \end{array}\right)&=h_\generation(b_n.\trunk.\delta.\challenge.\sigma\append\plot.\pi)\\
    \left(\begin{array}{c}
      \text{consensus rule~\ref{prop:consensus:answer}}\\
      \text{of Def.~\ref{def:blockchain}}
    \end{array}\right)&=h_\generation(\challenge_\mynext(b_{n-1}).\sigma\append\plot.\pi)\\
    \left(\begin{array}{c}
      \text{inductive}\\
      \text{hypothesis}
    \end{array}\right)&=h_\generation(\challenge_\mynext(b'_{n-1}).\sigma\append\plot.\pi)\\
    \left(\begin{array}{c}
      \text{consensus rule~\ref{prop:consensus:answer}}\\
      \text{of Def.~\ref{def:blockchain}}
    \end{array}\right)&=h_\generation(b'_n.\trunk.\delta.\challenge.\sigma\append\plot.\pi)\\
    \left(\begin{array}{c}
      \text{$b'_n$ has been}\\
      \text{mined by using}\\
      \text{$\plot$ only,}\\
      \text{and Def.~\ref{def:deadline_from_nonce}}
    \end{array}\right)&=h_\generation(b'_n.\trunk.\delta.\challenge.\sigma\append b'_n.\trunk.\delta.\pi)\\
    &=\sigma'.
  \end{align*}
  %
  It follows that $\challenge_\mynext(b_n)=\challenge_\mynext(b'_n)$.
  \qed
